### 2021-07-26 response.tex

\rev{Reviewer \#1}

\revtext{This is a review of ``Roles of generation-interval distributions in shaping relative epidemic strength, speed, and control of new SARS-CoV-2 variants'' by Park and others for Royal Society Interface.  This is the latest in an ongoing series of papers (Champredon 2015, Champredon 2018, Park 2019, Dushoff 2021, ...) in which the last author has considered aspects of incubation, growth, and reproduction in SIR epidemic theory.  In this manuscript, the authors use theory to anticipate the effects of COVID-19 variants where the distribution of infectivity evolves.}

\revtext{The manuscript is excellently written. The figures are clear.  All the mathematics seems satisfactory.  Unfortunately, the manuscript does not make its case for novel results.}

Thank you for the review.
We have tried to frame our results better, and to clarify both the novelty and importance of this study.
The primary purpose of our study is two-fold: (1) to illustrate the impact of uncertainty in the generation interval in shaping the population-level transmission advantage of new variants and (2) to evaluate when it is approprite to hold relative strength or speed constant.
While several studies of new variants have assumed constant relative speed (e.g., by fitting a logistic growth model), we argue that holding relative strength fixed is generally more appropriate given that the dynamics of SARS-CoV-2 have been driven primarily by strength-like interventions.
We also show that this relationship changes under speed-like interventions---that is, relative strength changes but relative speed remains constant.
We believe that the addition of a new simulations (Figure 5) and the revised text better highlight the novelty of our work.

### 2021-07-30 newvariant.tex
minor edits

Inferring the relative strength (i.e., the ratio of reproduction numbers, $\rho=\Rv/\Rw$) and relative speed (i.e., the difference between growth rates, $\delta=\rv-\rw$) of new SARS-CoV-2 variants compared to their wild types is critical to predicting and controlling the course of the current pandemic.
We explore how differences in the generation-interval distributions affect the inference of relative strength $\rho$ and speed $\delta$ in light of recent evidence that the alpha variant may have a longer infectious period (and therefore, a longer generation interval) than prior dominant lineages.
Neglecting these differences gives biased estimates of $\rho$, which can vary systematically with changes in the incidence of infections caused by the variant.
We further argue that holding $\rho$ constant is generally more appropriate in the context of SARS-CoV-2 and illustrate how $\rho$ and $\delta$ are affected by two idealized interventions.
A constant-strength intervention, which reduces transmission rates by a constant amount, affects $\delta$ but not $\rho$---this is analogous to vaccination and social distancing.
A constant-speed intervention, which isolates infected individuals at a constant hazard, affects $\rho$ but not $\delta$---this is analogous to contact tracing.
Uncertainties in the generation-interval distribution determine the effectiveness of these interventions.

@@

For example, when epidemic strength $\RR$ is fixed, assuming longer generation intervals leads to a slower epidemic growth (lower $r$), making the epidemic look easier to control with a constant-speed intervention, which isolates infected individuals at a constant hazard and therefore affect late transmission disproportionately---as shown in \cite{doi:10.1098/rspb.2020.1556}, the hazard of isolation must be greater than the epidemic speed $r$ to reach the $r=0$ threshold.
In this same case, however, the uncertainty in the generation-interval distribution is expected to have no effects on the effectiveness of a constant-strength intervention, which reduce the hazard of infectious contacts and therefore affect transmission by an equal amount regardless of age of infection.
Conversely, when epidemic speed $r$ is fixed, assuming longer generation intervals leads to a higher estimate of epidemic strength ($\RR$), making the epidemic look harder to control with a constant-strength interventions (but not with a constant-speed interventions, whose effectiveness is fixed if speed is fixed).
Constant strength and speed are idealized representations of real-life interventions, which can be either strength-like (e.g., vaccination and social distancing) or speed-like (e.g., contact tracing and isolation) depending on how their effectiveness varies across the generation interval (see \cite{doi:10.1098/rspb.2020.1556} and Discussion).

Most studies have assumed that the previous dominant strain (wild type) and variants have identical generation-interval distributions, but recent evidence suggests that the alpha variant may have a longer duration of infection: 13.3 days (90\% CI: 10.1--16.5) for the new variant and 8.2 days (90\% CI: 6.5--9.7) for the wild type \citep{kissler2021densely}.
Longer duration of infection suggests that the mean generation interval of the alpha variant is likely to be longer than that of the wild type \citep{lloyd2001realistic,wearing2005appropriate,roberts2007model}.
Other studies have considered the possibility that the faster growth rate of new variants may be driven, in part, by shorter generation intervals \citep{davies2021estimated,volz2021transmission}.
However, linking strength and speed is complicated given that the generation-interval distribution depend on many factors including behavior: 
for example, self-isolation after symptom onset will lead to shorter generation intervals.

@@

Neglecting the (relatively slow) rate of new mutations and assuming a homogeneous mixing, the current incidence of infection $i_x(t)$ caused by each strain $x$---either the wild type (``wt'') or the variant (``var'')---can be expressed in terms of their previous incidence $i_x(t-\tau)$ and the rate at which secondary infections are generated at time $t$ by individuals infected $\tau$ time units ago $K_x(t, \tau)$:
\begin{equation}
i_x(t) = \int_0^\infty i_x(t-\tau) K_x(t, \tau) \dtau.
\end{equation}
This framework provides a flexible way of modeling disease dynamics and generalizes a wide range of compartmental models, including the SEIR model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral,breda2012formulation, champredon2018equivalence}.

@@

Under constant-strength changes that reduce the transmission rate by a constant amount regardless of the age of infection, the instantaneous generation-interval distribution does not across change time \citep{fraser2007estimating}.
In this case, the instantaneous generation-interval distribution is also often referred to as the intrinsic generation-interval distribution \citep{champredon2015intrinsic,champredon2018two,gostic2020practical,park2020time}---
for example, the standard SEIR model can be equivalently expressed as a renewal equation with time-invariant intrinsic generation-interval distribution $g(\tau)$ as shown in \citep{champredon2018equivalence}.
The instantaneous generation-interval distribution is also different from the realized generation-interval distribution, which measures time between actual infection events (see \citep{champredon2015intrinsic}).
Previous studies have noted, in many contexts, that the realized generation intervals can contract due to susceptible depletion---a special case of constant-strength changes \citep{kenah2008generation,nishiura2010time,champredon2015intrinsic}---even though the instantaneous generation-interval distribution remains unchanged in this scenario.
While the instantaneous generation-interval distribution can change under speed-like changes (see \cite{fraser2007estimating} and Section 6--7 for detailed discussions), assuming a time-invariant instantaneous generation-interval distribution is often appropriate in the context of SARS-CoV-2, given that the dynamics of its spread have been primarily driven by strength-like intervention, such as social distancing measures \citep{flaxman2020Rt} and vaccination \citep{moore2021vaccination}.
Indeed, many dynamical models of SARS-CoV-2 infections have solely relied on constant-strength changes, either implicitly or explicitly assuming a time-invariant intrinsic generation-interval distribution (e.g., \citep{flaxman2020Rt,gostic2020practical,brauner2021inferring}).
Therefore, we neglect changes in the intrinsic generation-interval distribution over time for now and focus on the impact of constant-strength changes on the inference of dynamics of new SARS-CoV-2 variants.
We revisit these ideas in Section 6 and compare the effects of constant-speed interventions with those of constant-strength interventions.

Over a short period of time, we can assume that epidemiological conditions remain roughly constant: $\RR_x(t) \approx \RR_x$ and $g_x(t, \tau) \approx g_x(\tau)$, in which case the incidence of infections changes exponentially.
Here, we use the term ``epidemiological conditions'' to broadly refer to all factors that affect transmission---mathematically, they are captured by the kernel $K(t, \tau)$.
In the context of SARS-CoV-2 infections, we are essentially assuming that the changes in susceptible pool, behavior, and contact rates are usually small over a short period of time.
Then, the incidence of each strain grows (or decays) exponentially at rate $r_x$, satisfying the Euler-Lotka equation \citep{wallinga2007generation}:
\begin{equation}
\frac{1}{\RR_x} = \int_0^\infty \exp(- r_x \tau) g_x(\tau) \dtau.
\end{equation}
We can approximate this $r$--$\RR$ relationship by assuming that the generation-interval distribution is Gamma-distributed, and summarizing it using the mean generation interval $\bar{G}_x$ and the squared coefficient of variation $\kappa_x$ \citep{park2019practical}:
\begin{equation}
\RR_x \approx (1 + \kappa_x r_x \bar{G}_x)^{1/\kappa_x}.
\end{equation}
Various Gamma assumptions have been widely used in epidemic modeling, including for models of SARS-CoV-2 \citep{doi:10.1098/rsif.2020.0144}.
Although the Gamma assumption includes as a special case models that assume exponentially distributed generation intervals (when $\kappa=1$), corresponding to the SIR model \citep{anderson1991infectious}, it is different from the generation-interval distribution we get when infectious periods are Gamma distributed, another standard assumption in epidemic modeling (see \cite{roberts2007model} for discussion).

We use this framework to investigate how inferences about strength and speed of the variant depend on our assumptions about the underlying generation-interval distributions.
Here, we focus on differences in mean generation intervals, assuming that both the squared coefficient of variation and distribution shape of the variant are otherwise the same as for the wild type: $\kappa_{\mathrm{wt}} = \kappa_{\mathrm{var}} = \kappa$.
Changes in shape can be important \citep{miller2010epidemics,svensson2015influence}, and we do not investigate them here. 
We do note that, loosely, we expect a distribution with higher coefficient of variation to allow for more early transmission, and thus to have qualitatively similar effects to a distribution with a shorter mean in many cases \citep{park2019practical}.

@@

We start by taking the mean generation interval of the wild type to be $\Gw = 5\days$ \citep{ferretti2020quantifying}; and the squared coefficient of variation of generation intervals to be $\kappa=0.2$ \citep{ferretti2020quantifying} for both the variant and the wild type.
As noted above, we assume throughout that the variant and the wild type can be approximated with gamma distributions with equal $\kappa$, and only consider differences in the mean.
We evaluate the estimates of relative strength $\rho$ across a wide range of $\kappa$ from 0 (fixed-length generation intervals) to 1 (exponential distribution).
To further explore how inference depends on underlying epidemiological conditions, we consider five scenarios: (1) $\rw < \rv < 0$, (2) $\rw < \rv = 0$, (3) $\rw < 0 < \rv$, (4) $0 = \rw < \rv$, and (5) $0 < \rw < \rv$.
In all of these cases, we assume $\delta = \rv - \rw = 0.1\pday$ based on the observed relative growth rate of the alpha variant in the UK \citep{davies2021estimated}.

@@

Instead of inferring relative strength from speed, one can directly estimate time-varying reproduction numbers $\RR(t)$ of the variant and the wild type from incidence data and directly compare their ratios when the incidence of infection and the generation-interval distributions are known---such methods have been used in previous analyses of the alpha variant by \cite{volz2021transmission}.
Broadly, there are two types of time-varying reproduction numbers: case reproduction number and instantaneous reproduction number.
The case reproduction number is defined as the average number of secondary infections caused by an individual infected at time $t$ and therefore depends on transmission after time $t$ \citep{wallinga2004different}.
The instantaneous reproduction number is defined as the average number of secondary infections that would be caused by an individual infected at time $t$ if conditions were to remain the same \citep{fraser2007estimating}; 
therefore, the instantaneous reproduction number only depends on transmission at time $t$ and is most appropriate for real-time evaluation of changes in transmission \citep{gostic2020practical}.
The estimation of the instantaneous reproduction number was popularized by \cite{cori2013new} and has been widely adopted in epidemiological analyses of SARS-CoV-2 \citep{abbott2020estimating,knight2020estimating,flaxman2020Rt,brauner2021inferring,li2021temporal}.

@@

We consider two types of intervention:
an intervention of constant strength, which reduces transmission by a constant factor $\theta$ regardless of age of infection ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$); and an intervention of constant speed, which reduces transmission after infection by a constant rate $\phi$ ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$).
In this case, we can control the spread of the variant when $\theta > \Rv$ or $\phi > \rv$, respectively \citep{doi:10.1098/rspb.2020.1556}.
We consider constant-strength and -speed interventions calibrated to reduce $\Rv$ to 0.9 under the assumption that the variant generation intervals match those of the wild type(\fref{strengthspeed}A--B).
While both interventions are equally effective on the strength scale (that is, $\Ry{post}=\int  K_{\mathrm{post}}(\tau) \dtau = 0.9$), they have different dynamical implications.
The constant-strength intervention affects transmission equally throughout the course of infection, whereas the constant-speed intervention has greater impact on transmission that occurs later in infection;
as a result, the constant-speed intervention reduces the post-intervention mean generation interval (\fref{strengthspeed}B) and leads to (slightly) faster exponential decay (therefore, lower $\ry{post}$).

\begin{figure}[!pth]
\begin{center}
\includegraphics[width=0.95\textwidth]{control.pdf}
\caption{
\textbf{Effects of constant-strength and constant-speed interventions on the spread of a new variant with known speed \rv.}
(A--D) Pre-intervention (black) and post-intervention (colored) kernel of the new variant under constant-strength (A,C) and -speed (B,D) interventions when the variant has equal mean generation interval as the wild type (A--B) and when the variant has longer generation intervals than the wild type (C--D).
Vertical lines represent the mean generation interval.
(E) Pre-intervention $\Rv$ (black) and post-intervention epidemic strength $\Ry{post}$ (colored) conditional on the mean generation interval of the variant.
(F) Pre-intervention $\rv$ (black) and post-intervention epidemic speed $\ry{post}$ (colored) conditional on the mean generation interval of the variant.
Epidemic strength and speed are calculated assuming $\rw=0\pday$, $\delta=0.1\pday$, $\rv=0.1\pday$, $\Gy{wt}=5\,\textrm{days}$, and $\kappa=1/5$ for pre-intervention conditions.
The constant-strength intervention assumes that the ratio $\Rv/\Ry{post} = \theta$ remains constant.
The constant-speed intervention assumes that the difference $\rv - \ry{post} = \phi$ remains constant.
Intervention strength and speed are chosen so that post intervention strength of the new variant is 0.9 when its mean generation interval is 5 days.
}
\label{fig:strengthspeed}
\end{center}
\end{figure}

However, if the variant has longer generation intervals than the wild type (\fref{strengthspeed}C--D), then the strength of the variant will be higher because we are holding the observed speed constant (\fref{strengthspeed}E).
In this case, the same constant-strength intervention can fail to control the epidemic (i.e., $\Ry{post} > 1$; \fref{strengthspeed}E) because this intervention reduces the transmission by a constant amount regardless of age of infection (\fref{strengthspeed}C).
On the other hand, the same constant-speed intervention will prevent a larger proportion of transmission, leading to lower $\Ry{post}$ (\fref{strengthspeed}E), because it is more effective against late-stage transmission (\fref{strengthspeed}D).
The constant-speed intervention also reduces the mean generation interval by a larger factor (\fref{strengthspeed}D).

@@

Finally, we synthesize our findings and illustrate the differences between constant-strength and constant-speed interventions using epidemic simulations (\fref{control_sim}).
As before, we consider a scenario in which a new variant is emerging with known relative speed($\delta=\rv > 0.1\pday$ and $\Rw > 1$) while the incidence of infections caused by the wild type is constant ($\rw = 0\pday$ and $\Rw =1$) before interventions are reduced.
The constant-strength intervention then reduces both $\Rw$ and $\Rv$ by a constant factor $\theta$ beginning day 30, 
whereas the constant-speed intervention isolates individuals infected with both the wild type and the variant at a constant hazard $\phi$ after day 30.
If the variant has a longer mean generation interval, the constant-strength intervention fails to suppress the spread of the variant (\fref{control_sim}A) because the longer generations imply a higher initial $\Rw$, and thus a stronger intervention is required to reduce $\Rw$ below 1 (\fref{control_sim}B). 
The relative strength remains constant under constant-strength intervention (\fref{control_sim}B), but the relative speed changes (\fref{control_sim}C) as discussed earlier in Section 4---in this case, the longer mean generation interval variant increases the relative speed when the intervention is introduced.

In contrast, epidemic trajectories under the constant-speed intervention look identical regardless of the mean generation interval of the variant (\fref{control_sim}D). 
As shown in \fref{strengthspeed}E, a constant-speed intervention leads to greater reduction in $\Rv(t)$ when the mean generation-interval of the variant is longer (\fref{control_sim}E)---
this implies a change in relative strength $\rho$ because the effectiveness of a constant-speed intervention is sensitive to the generation-interval distribution.
In this case, the relative speed of the variant remains unaffected by the constant-speed intervention because the epidemic speed of the wild type $\rw$ and the variant $\rv$ are both reduced by equal amounts $\phi$, resulting in the post-intervention epidemic speed of $\rw-\phi$ and $\rv-\phi$, respectively.

@@

There is currently a (minor) discrepancy in the estimates of relative strength of new variants, particularly for the alpha variant.
Mathematical analyses have typically reported greater than a 1.4 fold increase in reproduction number for the alpha variant ($\rho > 1.4$) whereas an independent analysis of secondary attack rates from contact tracing data suggests a 1.25--1.4 fold increase \citep{ukinvest}.
As shown in \cite{davies2021estimated} and \cite{volz2021transmission}, propagating uncertainty in generation interval estimates (and, in particular, assuming shorter generation intervals) can partially explain these differences:
For example, if we consider short and wide generation interval estimates from Tianjin, China ($\Gw=\Gv=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.26$ (from $\delta=0.1\pday$ and $\rw = 0\pday$).
Although these estimates are more consistent with the attack rate analysis \citep{ukinvest},
we do not claim that they are necessarily more accurate---
while individual-level data from contact tracing can provide a more reliable source of information about the transmissibility and time scale of transmission in some cases, they can also be biased towards particular types of contacts---for example, household contacts are probably more likely to be identified---which could also affect the estimate of $\rho$.
Instead, this calculation simply highlights the importance of propagating uncertainty in generation-interval distributions in assessing the relative strength and speed of SARS-CoV-2 variants.

>>>>

Inferring the relative strength (i.e., the ratio of reproduction numbers) and relative speed (i.e., the difference between growth rates) of new SARS-CoV-2 variants compared to their wild types is critical to predicting and controlling the course of the current pandemic.
We explore how differences in the generation-interval distributions affect the inference of relative strength and speed in light of recent evidence that the Alpha and Delta variants may have different infectiousness profiles (and therefore, the generation interval) from prior dominant lineages.
Neglecting these differences gives biased estimates of relative strength, which can vary systematically with changes in the incidence of infections caused by the variant.
We further argue that holding relative strength constant is generally more appropriate in the context of SARS-CoV-2 and illustrate how relative strength and speed are affected by two idealized interventions.
A constant-strength intervention, which reduces transmission rates by a constant amount, affects relative speed but not relative strength---this is analogous to vaccination and social distancing.
A constant-speed intervention, which isolates infected individuals at a constant hazard, affects relative strength but not relative speed---this is analogous to contact tracing.
Changes in the generation-interval distribution can alter the effectiveness of these interventions.

@@

For example, when epidemic strength $\RR$ is fixed, assuming longer generation intervals leads to a slower epidemic growth (lower $r$), making the epidemic look easier to control with a constant-speed intervention, which isolates infected individuals at a constant hazard and therefore affects late transmission disproportionately.
As shown in \cite{doi:10.1098/rspb.2020.1556}, the hazard of isolation must be greater than the epidemic speed $r$ to reach the $r=0$ threshold.
In this same case, however, the uncertainty in the generation-interval distribution is expected to have no effects on the effectiveness of a constant-strength intervention, which reduces the hazard of infectious contacts and therefore affects transmission by an equal amount regardless of age of infection.
Conversely, when epidemic speed $r$ is fixed, assuming longer generation intervals leads to a higher estimate of epidemic strength ($\RR$), making the epidemic look harder to control with a constant-strength interventions.
Constant strength and speed are idealized representations of real-life interventions, which can be either strength-like (e.g., vaccination and social distancing) or speed-like (e.g., contact tracing and isolation) depending on how their effectiveness varies across the generation interval (see \cite{doi:10.1098/rspb.2020.1556} and Discussion).

Most studies have assumed that the previous dominant strain (wild type) and variants have identical generation-interval distributions, but Kissler \textit{et al.} \citep{kissler2021densely} suggested that the Alpha variant may have a longer duration of infection: 13.3 days (90\% CI: 10.1--16.5) for the new variant and 8.2 days (90\% CI: 6.5--9.7) for the wild type.
Longer duration of infection suggests that the mean generation interval of the Alpha variant is likely to be longer than that of the wild type \citep{lloyd2001realistic,wearing2005appropriate,roberts2007model}.
In contrast, some studies have suggested that the more recent Delta variant may have a shorter generation interval \citep{li2021viral, zhang2021transmission}.
Modeling studies have also considered the possibility that the faster growth rate of new variants may be driven, in part, by shorter generation intervals \citep{davies2021estimated,volz2021transmission}.
However, linking strength and speed is complicated given that the generation-interval distribution depend on many factors including behavior: 
for example, self-isolation after symptom onset will lead to shorter generation intervals.

@@

Neglecting the (relatively slow) rate of new mutations and assuming homogeneous mixing, the current incidence of infection $i_x(t)$ caused by each strain $x$---either the wild type (``wt'') or the variant (``var'')---can be expressed in terms of their previous incidence $i_x(t-\tau)$ and the rate at which secondary infections are generated at time $t$ by individuals infected $\tau$ time units ago $K_x(t, \tau)$:
\begin{equation}
i_x(t) = \int_0^\infty i_x(t-\tau) K_x(t, \tau) \dtau.
\end{equation}
This framework provides a flexible way of modeling disease dynamics and generalizes a wide range of compartmental models, including the SEIR model \citep{heesterbeek1996concept, diekmann2000mathematical, roberts2004modelling, aldis2005integral,breda2012formulation, champredon2018equivalence}.

@@

Under constant-strength changes that reduce the transmission rate by a constant amount regardless of the age of infection, the instantaneous generation-interval distribution does not change across time \citep{fraser2007estimating}.
In this case, the instantaneous generation-interval distribution is also often referred to as the intrinsic generation-interval distribution \citep{champredon2015intrinsic,champredon2018two,gostic2020practical,park2020time}---
for example, the standard SEIR model can be equivalently expressed as a renewal equation with time-invariant intrinsic generation-interval distribution $g(\tau)$ as shown in \citep{champredon2018equivalence}.
The instantaneous generation-interval distribution is also different from the realized generation-interval distribution, which measures time between actual infection events (see \citep{champredon2015intrinsic}).
Previous studies have noted, in many contexts, that the realized generation intervals can contract due to susceptible depletion---a special case of constant-strength changes \citep{kenah2008generation,nishiura2010time,champredon2015intrinsic}---even though the instantaneous generation-interval distribution remains unchanged in this scenario.
While the instantaneous generation-interval distribution can change under speed-like changes (see \cite{fraser2007estimating} and Section 6--7 for detailed discussions), assuming a time-invariant instantaneous generation-interval distribution is often appropriate in the context of SARS-CoV-2, given that the dynamics of its spread have been primarily driven by strength-like intervention, such as social distancing measures \citep{flaxman2020Rt} and vaccination \citep{moore2021vaccination}.
Indeed, many dynamical models of SARS-CoV-2 infections have solely relied on constant-strength changes, either implicitly or explicitly assuming a time-invariant intrinsic generation-interval distribution (e.g., \citep{flaxman2020Rt,gostic2020practical,brauner2021inferring}).
Therefore, we neglect changes in the intrinsic generation-interval distribution over time for now and focus on the impact of constant-strength changes on the inference of dynamics of new SARS-CoV-2 variants.
We revisit these ideas in Section 6 and compare the effects of constant-speed interventions with those of constant-strength interventions.

Over a short period of time, we can assume that epidemiological conditions remain roughly constant: $\RR_x(t) \approx \RR_x$ and $g_x(t, \tau) \approx g_x(\tau)$, in which case the incidence of infections changes exponentially.
Here, we use the term ``epidemiological conditions'' to broadly refer to all factors that affect transmission---mathematically, they are captured by the kernel $K(t, \tau)$.
In the context of SARS-CoV-2 infections, we are essentially assuming that the changes in susceptible pool, behavior, and contact rates are usually small over a short period of time.
Then, the incidence of each strain grows (or decays) exponentially at rate $r_x$, satisfying the Euler-Lotka equation \citep{wallinga2007generation}:
\begin{equation}
\frac{1}{\RR_x} = \int_0^\infty \exp(- r_x \tau) g_x(\tau) \dtau.
\end{equation}
We can approximate this $r$--$\RR$ relationship by assuming that the generation-interval distribution is Gamma-distributed, and summarizing it using the mean generation interval $\bar{G}_x$ and the squared coefficient of variation $\kappa_x$ \citep{park2019practical}:
\begin{equation}
\RR_x \approx (1 + \kappa_x r_x \bar{G}_x)^{1/\kappa_x}.
\end{equation}
Various Gamma-generation-interval assumptions have been widely used in epidemic modeling, including for models of SARS-CoV-2 \citep{doi:10.1098/rsif.2020.0144}.
The Gamma-generation-interval assumption includes as a special case models that assume exponentially distributed generation intervals (when $\kappa=1$), corresponding to the SIR model \citep{anderson1991infectious}.
We note that when the infectious periods are Gamma distributed---another standard assumption in epidemic modeling---the resulting generation interval does not follow the Gamma distribution;
in particular, the mean generation interval can be shorter than the mean infectious period because transmission occurs throughout the infectious period until recovery (see \cite{roberts2007model} for detailed discussion).

We use this framework to investigate how inferences about strength and speed of the variant depend on our assumptions about the underlying generation-interval distributions.
Here, we focus on differences in mean generation intervals, assuming that both the squared coefficient of variation and distribution shape of the variant are otherwise the same as for the wild type: $\kappa_{\mathrm{wt}} = \kappa_{\mathrm{var}} = \kappa$.
Changes in shape can be important \citep{miller2010epidemics,svensson2015influence}, and we do not investigate them here. 
We do note that we expect a distribution with higher coefficient of variation to allow for more early transmission, and thus to have qualitatively similar effects to a distribution with a shorter mean in many cases \citep{park2019practical}.

@@

We start by taking the mean generation interval of the wild type to be $\Gw = 5\days$ \citep{ferretti2020quantifying}; and the squared coefficient of variation of generation intervals to be $\kappa=0.2$ \citep{ferretti2020quantifying} for both the variant and the wild type.
As noted above, we assume throughout that the variant and the wild type can be approximated with gamma distributions with equal $\kappa$, and only consider differences in the mean.
We evaluate the estimates of relative strength $\rho$ across a wide range of $\kappa$ from 0 (fixed-length generation intervals) to 1 (exponential distribution).
To further explore how inference depends on underlying epidemiological conditions, we consider five scenarios: (1) $\rw < \rv < 0$, (2) $\rw < \rv = 0$, (3) $\rw < 0 < \rv$, (4) $0 = \rw < \rv$, and (5) $0 < \rw < \rv$.
In all of these cases, we assume $\delta = \rv - \rw = 0.1\pday$ based on the observed relative growth rate of the Alpha variant in the UK \citep{davies2021estimated}.

@@

Instead of inferring relative strength from speed, one can directly estimate time-varying reproduction numbers $\RR(t)$ of the variant and the wild type from incidence data and directly compare their ratios when the incidence of infection and the generation-interval distributions are known---such methods have been used in previous analyses of the Alpha variant by \cite{volz2021transmission}.
Broadly, there are two types of time-varying reproduction numbers: case reproduction number and instantaneous reproduction number.
The case reproduction number is defined as the average number of secondary infections caused by an individual infected at time $t$ and therefore depends on transmission after time $t$ \citep{wallinga2004different}.
The instantaneous reproduction number is defined as the average number of secondary infections that would be caused by an individual infected at time $t$ if conditions were to remain the same \citep{fraser2007estimating}; 
therefore, the instantaneous reproduction number only depends on transmission at time $t$ and is most appropriate for real-time evaluation of changes in transmission \citep{gostic2020practical}.
The estimation of the instantaneous reproduction number was popularized by \cite{cori2013new} and has been widely adopted in epidemiological analyses of SARS-CoV-2 \citep{abbott2020estimating,knight2020estimating,flaxman2020Rt,brauner2021inferring,li2021temporal}.

@@

We consider two types of intervention:
an intervention of constant strength (\fref{strengthspeed}A,C,E), which reduces transmission by a constant factor $\theta$ regardless of age of infection ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau)/\theta$); and an intervention of constant speed (\fref{strengthspeed}B,D,F), which reduces transmission after infection by a constant rate $\phi$ ($K_{\mathrm{post}}(\tau) = K_{\mathrm{pre}}(\tau) \exp(-\phi \tau)$).
In this case, we can control the spread of the variant when $\theta > \Rv$ or $\phi > \rv$, respectively \citep{doi:10.1098/rspb.2020.1556}.
We consider constant-strength and -speed interventions calibrated to reduce $\Rv$ to 0.9 under the assumption that the variant generation intervals match those of the wild type (\fref{strengthspeed}C--D).
While both interventions are equally effective on the strength scale (that is, $\Ry{post}=\int  K_{\mathrm{post}}(\tau) \dtau = 0.9$), they have different dynamical implications.
The constant-strength intervention affects transmission equally throughout the course of infection, whereas the constant-speed intervention has greater impact on transmission that occurs later in infection;
as a result, the constant-speed intervention reduces the post-intervention mean generation interval (\fref{strengthspeed}B) and leads to (slightly) faster exponential decay (therefore, lower $\ry{post}$).

\begin{figure}[!pth]
\begin{center}
\includegraphics[width=0.95\textwidth]{control.pdf}
\caption{
\textbf{Effects of constant-strength and constant-speed interventions on the spread of a new variant with known speed \rv.}
(A--F) Pre-intervention (black) and post-intervention (colored) kernel of the new variant under constant-strength (A,C,E) and -speed (B,D,F) interventions when the variant has shorter (A--B), equal (C--D), or longer (E--F) mean generation interval than the wild type.
Vertical lines represent the mean generation interval.
(G) Pre-intervention $\Rv$ (black) and post-intervention epidemic strength $\Ry{post}$ (colored) conditional on the mean generation interval of the variant.
(H) Pre-intervention $\rv$ (black) and post-intervention epidemic speed $\ry{post}$ (colored) conditional on the mean generation interval of the variant.
Epidemic strength and speed are calculated assuming $\rw=0\pday$, $\delta=0.1\pday$, $\rv=0.1\pday$, $\Gy{wt}=5\,\textrm{days}$, and $\kappa=1/5$ for pre-intervention conditions.
The constant-strength intervention assumes that the ratio $\Rv/\Ry{post} = \theta$ remains constant.
The constant-speed intervention assumes that the difference $\rv - \ry{post} = \phi$ remains constant.
Intervention strength and speed are chosen so that post intervention strength of the new variant is 0.9 when its mean generation interval is 5 days.
}
\label{fig:strengthspeed}
\end{center}
\end{figure}

However, if the variant has longer generation intervals than the wild type (\fref{strengthspeed}E--F), then the strength of the variant will be higher because we are holding the observed speed constant (\fref{strengthspeed}G).
In this case, the same constant-strength intervention can fail to control the epidemic (i.e., $\Ry{post} > 1$; \fref{strengthspeed}H) because this intervention reduces the transmission by a constant amount regardless of age of infection (\fref{strengthspeed}E).
On the other hand, the same constant-speed intervention will prevent a larger proportion of transmission, leading to lower $\Ry{post}$ (\fref{strengthspeed}G), because it is more effective against late-stage transmission (\fref{strengthspeed}F).
The constant-speed intervention also reduces the mean generation interval by a larger factor (\fref{strengthspeed}F).
Conversely, if the variant has shorter generation intervals than the wild type (\fref{strengthspeed}A--B), the given constant-strength intervention will be relative more effective (\fref{strengthspeed}A) because the given constant-speed intervention prevents less transmission (\fref{strengthspeed}B).

@@

Finally, we synthesize our findings and illustrate the differences between constant-strength and constant-speed interventions using epidemic simulations (\fref{control_sim}).
As before, we consider a scenario in which a new variant is emerging with known relative speed($\delta=\rv > 0.1\pday$ and $\Rw > 1$) while the incidence of infections caused by the wild type is constant ($\rw = 0\pday$ and $\Rw =1$) before interventions are reduced.
The constant-strength intervention then reduces both $\Rw$ and $\Rv$ by a constant factor $\theta$ beginning day 30, 
whereas the constant-speed intervention isolates individuals infected with both the wild type and the variant at a constant hazard $\phi$ after day 30.
If the variant has a longer mean generation interval, this specific constant-strength intervention fails to suppress the spread of the variant (\fref{control_sim}A) because the longer generations imply a higher initial $\Rw$, and thus a stronger intervention is required to reduce $\Rw$ below 1 (\fref{control_sim}B).
Conversely, if the variant has a shorter mean generation interval, its initial $\Rw$ will be lower, and therefore the same constant-strength intervention will be more effective, leading to lower post-intervention $\Rw$ (\fref{control_sim}B).
The relative strength remains constant under constant-strength intervention (\fref{control_sim}B), but the relative speed changes (\fref{control_sim}C) as discussed earlier in Section 4---in this case, the longer mean generation interval variant increases the relative speed when the intervention is introduced.

In contrast, epidemic trajectories under the constant-speed intervention look identical regardless of the mean generation interval of the variant (\fref{control_sim}D). 
As shown in \fref{strengthspeed}E, a constant-speed intervention leads to greater reduction in $\Rv(t)$ when the mean generation-interval of the variant is longer (\fref{control_sim}E)---
this implies a change in relative strength $\rho$ because the resulting post-intervention strength under a constant-speed intervention is sensitive to the generation-interval distribution as shown in \fref{strengthspeed}.
In this case, the relative speed of the variant remains unaffected by the constant-speed intervention because the epidemic speed of the wild type $\rw$ and the variant $\rv$ are both reduced by equal amounts $\phi$, resulting in the post-intervention epidemic speed of $\rw-\phi$ and $\rv-\phi$, respectively.

@@

There is currently a (minor) discrepancy in the estimates of relative strength of new variants, particularly for the Alpha variant.
Mathematical analyses have typically reported greater than a 1.4 fold increase in reproduction number for the Alpha variant ($\rho > 1.4$) whereas an independent analysis of secondary attack rates from contact tracing data suggests a 1.25--1.4 fold increase \citep{ukinvest}.
As shown in \cite{davies2021estimated} and \cite{volz2021transmission}, propagating uncertainty in generation interval estimates (and, in particular, assuming shorter generation intervals) can partially explain these differences:
For example, if we consider short and wide generation interval estimates from Tianjin, China ($\Gw=\Gv=2.57\days$ and $\kappa=1$; \cite{ganyani2020estimating}), we obtain $\rho=1.26$ (from $\delta=0.1\pday$ and $\rw = 0\pday$).
Although these estimates are more consistent with the attack rate analysis \citep{ukinvest},
we do not claim that they are necessarily more accurate---
while individual-level data from contact tracing can provide a more reliable source of information about the transmissibility and time scale of transmission in some cases, they can also be biased towards particular types of contacts---for example, household contacts are probably more likely to be identified---which could also affect the estimate of $\rho$.
Instead, this calculation simply highlights the importance of propagating uncertainty in generation-interval distributions in assessing the relative strength and speed of SARS-CoV-2 variants.
While we relied on parameter estimates for the Alpha variant throughout our analysis, our qualitative conclusions can be applied to studying other SARS-CoV-2 variants, including the Delta variant.
